A comparison of classification algorithms applied to the Adult Database

### 2009-12-16 demo/demo.tex
general latex syntax fixes.

    \begin{itemize}
    \item<1-> Given an adult with some attributes, identify whether he gains more than 50000 dollars per year, or not.
    \item<1-> Identify what social characteristics allow one person to gain more than someone else.
    \item<1-> Understanding what affects the income can give a lot of information concerning society's behaviour.
    \end{itemize}
}

@@

    \begin{itemize}
    \item<1-> 48842 entries of individual information of people gathered in a 1994 census on the United States of America.
    \begin{itemize}
        \item<1-> 32561 records for the training set (67\%).
        \item<1-> 16281 records for the test set (33\%).
    \end{itemize}
    \item<1-> Each entry is composed by a set of 14 attributes and 2 classes (winning more or less than 50K dollars a year).
    \item<1-> There are 3620 records with missing values. The only attributes containing missing values are \textbf{Workclass}, \textbf{Occupation} and \textbf{Native-country}.
    \item<1-> Available in \url{ftp://ftp.ics.uci.edu/pub/machine-learning-databases/adult}.
    \end{itemize}
}

@@

    \begin{itemize}
        \item<1-> \textbf{Age} (continuous).
        \item<1-> \textbf{Workclass} (categorical).
        \item<1-> \textbf{FNLWGT} (continuous).
        \item<1-> \textbf{Education} (ordinal).
        \item<1-> \textbf{Education-num} (continuous).
        \item<1-> \textbf{Marital-status} (categorical).
        \item<1-> \textbf{Occupation} (categorical).
        \item<1-> \textbf{Relationship} (categorical).
        \item<1-> \textbf{Race} (categorical).
        \item<1-> \textbf{Sex} (categorical).
        \item<1-> \textbf{Capital-gain} (continuous).
        \item<1-> \textbf{Capital-loss} (continuous).
        \item<1-> \textbf{Hours-per-week} (continuous).
        \item<1-> \textbf{Native country} (categorical).
    \end{itemize}   
}

@@

    \begin{itemize}
    \item<1-> There is a higher percentage of older people gaining more than 50K than younger ones.
    \item<1-> There is a higher percentage of people with more education gaining more than 50K than people with less studies.
    \item<1-> There is a higher percentage of workaholics gaining more than 50K than people who work in part-time.
    \end{itemize}

@@

    \begin{itemize}
      \item <1-> Association rules can be used to discover interesting relations and patterns between attributes.

@@

    \begin{itemize}
    \item<1-> Clustering can be used to detect similar individuals and to build classes in the absence of them.
    \item<1-> Generating 6 clusters with only three variables (\textbf{age}, \textbf{education\_num} and \textbf{hours\_per\_week}) leads to some interesting observations:
    \begin{itemize}
        \item<1-> Younger people are grouped together.
        \item<1-> People with different levels of education belong to different clusters.
    \end{itemize}
    \end{itemize}
}

@@

    \begin{itemize}
    \item<1-> The database poses a classification problem.
    \item<1-> There is no single classifier that works best on all given problems:
    \begin{itemize}
        \item<1-> Decision trees.
        \item<1-> Rule induction.
        \item<1-> Inductive logic programming.
        \item<1-> Bayesian classifiers.
    \end{itemize}
    \end{itemize}
}

@@

  \begin{itemize}
    \item<1-> Decision trees

    \begin{itemize}
      \item<1-> ID3
      \item<1-> C4.5
      \item<1-> CART
      \item<1-> Random Forest
      \item<1-> Alternating Decision Trees (ADTree)
    \end{itemize}

    \item<1-> Rule Induction

    \begin{itemize}
      \item <1-> CN2
    \end{itemize}

    \item<1-> Inductive Logic Programming
    \begin{itemize}
        \item<1-> Aleph
    \end{itemize}

    \item<1-> Bayesian Networks. Using different search algorithms:

    \begin{itemize}
      \item<1-> Genetic Search
      \item<1-> Hill Climbing
      \item<1-> Simulated Annealing
    \end{itemize}

@@

  \begin{itemize}
    \item<1-> Classifiers were built using the train set and then tested against the test set
    \item<1-> C4.5, Random Forest, CART, ADTree and Bayesian classifiers built using Weka
    \begin{itemize}
      \item<1-> Various parameters available in Weka algorithms were experimented in hope of getting better error rates
      \item<1-> Every missing value was replaced with modes and means
      \item<1-> Two different classifiers were trained for each algorithm, one using only discretized attributes
                and another with the attributes unmodified
      \item<1-> Analyze how the algorithm behaves when using numeric attributes
    \end{itemize}
  \end{itemize}
}

@@

  \begin{itemize}
    \item<1-> RapidMiner was used to build a classifier with the ID3 algorithm.
    \item<1-> Peter Clark's CN2 was used to induce rules.
    \begin{itemize}
        \item<1-> \url{http://www.cs.utexas.edu/users/pclark/software/\#cn2}.
    \end{itemize}
    \item<1-> Aleph was used to inductively generate logic clauses.
    \begin{itemize}
        \item<1-> \url{http://www.comlab.ox.ac.uk/activities/machinelearning/Aleph/aleph.html}.
    \end{itemize}
  \end{itemize}
  % colocar ILP ID3!
}

@@

  \begin{itemize}
    \item<1-> It is an ensemble classifier consisting of many decision trees
    \item<1-> Outputs the class that is the mode of the class's output by individual trees
    \item<1-> Better results require more time building the classifier
    \item<1-> Large quantities of memory needed: 1GB to compute with 400 trees
    \item<1-> Over-fitting
  \end{itemize}

@@

  \begin{itemize}
    \item<1-> Based on conditional probabilities and the Bayes theorem
    \item<1-> Represents a set of variables and their conditional independencies via a directed acyclic graph and a set of probability tables
    \item<1-> The algorithm first learns the network structure / graph and then learns the probability tables
    \item<1-> We experimented with different algorithms to learn the network structure:
     \begin{itemize}
      \item<1-> Genetic Search
      \item<1-> Hill Climbing
      \item<1-> Simulated Annealing
     \end{itemize}
    \item<1-> Weka provides a very rich set of parameters for each search algorithm
  \end{itemize}

@@

    \begin{itemize}
    \item<1-> CN2 can generate both ordered and unordered set of rules.
    \item<1-> Peter Clark's software can use different error estimations:
    \begin{itemize}
        \item<1-> Laplacian.
        \item<1-> Naive.
        \item<1-> Information Gain.
        \item<1-> Modified Entropy.
    \end{itemize}
    \end{itemize}

@@

    \begin{itemize}
    \item<1-> ID3 and C4.5 are very fast and build classifiers with a decent error rate.
    \item<1-> ILP and CN2 build more expressive classifiers, capable of a better interpretation by a human agent, but take longer to finish.
    \item<1-> The worst error rate is below 25\%.
    \item<1-> Using numeric information results in better error rates
    \item<1-> There is a lot of software tools that support the building of classifiers, namely Weka and RapidMiner.
    \begin{itemize}
        \item<1-> Open source.
        \item<1-> Easy to use.
        \item<1-> APIs for integration with Java applications.
    \end{itemize}
    \end{itemize}
}

@@

    \begin{itemize}
    \item<1-> Having a high capital gain clearly contributes to gain more than 50K dollars per year.
    \item<1-> Middle-aged husbands have a higher tendency to gain more than 50K dollars per year.
    \item<1-> A high percentage of people who gain less than 50K dollars per year are own childs.
    \item<1-> People coming from countries such as Taiwan, Thailand or Cambodia usually have higher education and gain more than 50K dollars per year.
    \end{itemize}
}

>>>>

    \begin{itemize}
    \item Given an adult with some attributes, identify whether he gains more than 50000 dollars per year, or not.
    \item Identify what social characteristics allow one person to gain more than someone else.
    \item Understanding what affects the income can give a lot of information concerning society's behaviour.
    \end{itemize}
}

@@

    \begin{itemize}
    \item 48842 entries of individual information of people gathered in a 1994 census on the United States of America.
    \begin{itemize}
        \item 32561 records for the training set (67\%).
        \item 16281 records for the test set (33\%).
    \end{itemize}
    \item Each entry is composed by a set of 14 attributes and 2 classes (winning more or less than 50K dollars a year).
    \item There are 3620 records with missing values. The only attributes containing missing values are \textbf{Workclass}, \textbf{Occupation} and \textbf{Native-country}.
    \item Available in \url{ftp://ftp.ics.uci.edu/pub/machine-learning-databases/adult}.
    \end{itemize}
}

@@

    \begin{itemize}
        \item \textbf{Age} (continuous).
        \item \textbf{Workclass} (categorical).
        \item \textbf{FNLWGT} (continuous).
        \item \textbf{Education} (ordinal).
        \item \textbf{Education-num} (continuous).
        \item \textbf{Marital-status} (categorical).
        \item \textbf{Occupation} (categorical).
        \item \textbf{Relationship} (categorical).
        \item \textbf{Race} (categorical).
        \item \textbf{Sex} (categorical).
        \item \textbf{Capital-gain} (continuous).
        \item \textbf{Capital-loss} (continuous).
        \item \textbf{Hours-per-week} (continuous).
        \item \textbf{Native country} (categorical).
    \end{itemize}   
}

@@

    \begin{itemize}
    \item There is a higher percentage of older people gaining more than 50K than younger ones.
    \item There is a higher percentage of people with more education gaining more than 50K than people with less studies.
    \item There is a higher percentage of workaholics gaining more than 50K than people who work in part-time.
    \end{itemize}

@@

    \begin{itemize}
      \item  Association rules can be used to discover interesting relations and patterns between attributes.

@@

    \begin{itemize}
    \item Clustering can be used to detect similar individuals and to build classes in the absence of them.
    \item Generating 6 clusters with only three variables (\textbf{age}, \textbf{education\_num} and \textbf{hours\_per\_week}) leads to some interesting observations:
    \begin{itemize}
        \item Younger people are grouped together.
        \item People with different levels of education belong to different clusters.
    \end{itemize}
    \end{itemize}
}

@@

    \begin{itemize}
    \item The database poses a classification problem.
    \item There is no single classifier that works best on all given problems:
    \begin{itemize}
        \item Decision trees.
        \item Rule induction.
        \item Inductive logic programming.
        \item Bayesian classifiers.
    \end{itemize}
    \end{itemize}
}

@@

  \begin{itemize}
    \item Decision trees

    \begin{itemize}
      \item ID3
      \item C4.5
      \item CART
      \item Random Forest
      \item Alternating Decision Trees (ADTree)
    \end{itemize}

    \item Rule Induction

    \begin{itemize}
      \item  CN2
    \end{itemize}

    \item Inductive Logic Programming
    \begin{itemize}
        \item Aleph
    \end{itemize}

    \item Bayesian Networks. Using different search algorithms:

    \begin{itemize}
      \item Genetic Search
      \item Hill Climbing
      \item Simulated Annealing
    \end{itemize}

@@

  \begin{itemize}
    \item Classifiers were built using the train set and then tested against the test set
    \item C4.5, Random Forest, CART, ADTree and Bayesian classifiers built using Weka
    \begin{itemize}
      \item Various parameters available in Weka algorithms were experimented in hope of getting better error rates
      \item Every missing value was replaced with modes and means
      \item Two different classifiers were trained for each algorithm, one using only discretized attributes
                and another with the attributes unmodified
      \item Analyze how the algorithm behaves when using numeric attributes
    \end{itemize}
  \end{itemize}
}

@@

  \begin{itemize}
    \item RapidMiner was used to build a classifier with the ID3 algorithm.
    \item Peter Clark's CN2 was used to induce rules.
    \begin{itemize}
        \item \url{http://www.cs.utexas.edu/users/pclark/software/\#cn2}.
    \end{itemize}
    \item Aleph was used to inductively generate logic clauses.
    \begin{itemize}
        \item \url{http://www.comlab.ox.ac.uk/activities/machinelearning/Aleph/aleph.html}.
    \end{itemize}
  \end{itemize}
}

@@

  \begin{itemize}
    \item It is an ensemble classifier consisting of many decision trees
    \item Outputs the class that is the mode of the class's output by individual trees
    \item Better results require more time building the classifier
    \item Large quantities of memory needed: 1GB to compute with 400 trees
    \item Over-fitting
  \end{itemize}

@@

  \begin{itemize}
    \item Based on conditional probabilities and the Bayes theorem
    \item Represents a set of variables and their conditional independencies via a directed acyclic graph and a set of probability tables
    \item The algorithm first learns the network structure / graph and then learns the probability tables
    \item We experimented with different algorithms to learn the network structure:
     \begin{itemize}
      \item Genetic Search
      \item Hill Climbing
      \item Simulated Annealing
     \end{itemize}
    \item Weka provides a very rich set of parameters for each search algorithm
  \end{itemize}

@@

    \begin{itemize}
    \item CN2 can generate both ordered and unordered set of rules.
    \item Peter Clark's software can use different error estimations:
    \begin{itemize}
        \item Laplacian.
        \item Naive.
        \item Information Gain.
        \item Modified Entropy.
    \end{itemize}
    \end{itemize}

@@

    \begin{itemize}
    \item ID3 and C4.5 are very fast and build classifiers with a decent error rate.
    \item ILP and CN2 build more expressive classifiers, capable of a better interpretation by a human agent, but take longer to finish.
    \item The worst error rate is below 25\%.
    \item Using numeric information results in better error rates
    \item There is a lot of software tools that support the building of classifiers, namely Weka and RapidMiner.
    \begin{itemize}
        \item Open source.
        \item Easy to use.
        \item APIs for integration with Java applications.
    \end{itemize}
    \end{itemize}
}

@@

    \begin{itemize}
    \item Having a high capital gain clearly contributes to gain more than 50K dollars per year.
    \item Middle-aged husbands have a higher tendency to gain more than 50K dollars per year.
    \item A high percentage of people who gain less than 50K dollars per year are own childs.
    \item People coming from countries such as Taiwan, Thailand or Cambodia usually have higher education and gain more than 50K dollars per year.
    \end{itemize}
}
